---
title: "Homework 6"
author: "Tianyou Wang"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(viridis)
library(dplyr)
library(rnoaa)
library(modelr)
library(purrr)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 10, 
  fig.height = 6,
  out.width = "90%"
)
```



## Problem 1 with Given Code - NOAA

### *Data Import*

```{r weather_df, cache = TRUE}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
```

To obtain a distribution for $\hat{r}^2$, we'll follow basically the same procedure we used for regression coefficients: draw bootstrap samples --> fit the a model to each --> extract the value I'm concerned with --> and summarize. Here, we'll use `modelr::bootstrap` to draw 5000 bootstrap samples and `broom::glance` to produce `r.squared` values. 

### *Distribution of $\hat{r}^2$ and its 95% CI*

```{r}
weather_r2_df = 
  weather_df %>% 
  modelr::bootstrap(n = 5000) %>% 
  mutate(
    models = map(strap, ~lm(tmax ~ tmin, data = .x) ),
    results = map(models, broom::glance)) %>% 
  select(-strap, -models) %>% 
  unnest(results)

weather_r2_df %>% 
  ggplot(aes(x = r.squared)) + 
  geom_density()
```

In this example, the $\hat{r}^2$ value is high, and the upper bound at 1 may be a cause for the generally skewed shape of the distribution. If we wanted to construct a confidence interval for $R^2$, we could take the 2.5% and 97.5% quantiles of the estimates across bootstrap samples. However, because the shape isn't symmetric, using the mean +/- 1.96 times the standard error probably wouldn't work well.

```{r}
weather_r2_df %>% 
  summarize(
    ci_r2_lower = quantile(r.squared, 0.025), 
    ci_r2_upper = quantile(r.squared, 0.975)) %>% 
  knitr::kable(digits = 3)
```



### Distribution of $\log(\beta_0 * \beta1)$ and its 95% CI

We can produce a distribution for $\log(\beta_0 * \beta1)$ using a similar approach, with a bit more wrangling before we make our plot.

```{r}
weather_logb0b1_df = 
  weather_df %>% 
  modelr::bootstrap(n = 5000) %>% 
  mutate(
    models = map(strap, ~lm(tmax ~ tmin, data = .x) ),
    results = map(models, broom::tidy)) %>% 
  select(-strap, -models) %>% 
  unnest(results) %>% 
  select(id = `.id`, term, estimate) %>% 
  pivot_wider(
    names_from = term, 
    values_from = estimate) %>% 
  rename(beta0 = `(Intercept)`, beta1 = tmin) %>% 
  mutate(log_b0b1 = log(beta0 * beta1))


weather_logb0b1_df %>% 
  ggplot(aes(x = log_b0b1)) + 
  geom_density()


weather_logb0b1_df %>% 
  summarize(
    ci_logb0b1_lower = quantile(log_b0b1, 0.025), 
    ci_logb0b1_upper = quantile(log_b0b1, 0.975)) %>% 
  knitr::kable(digits = 3)
```

As with $r^2$, this distribution is somewhat skewed and has some outliers. 

The point of this is not to say you should always use the bootstrap -- it's possible to establish "large sample" distributions for strange parameters / values / summaries in a lot of cases, and those are great to have. But it is helpful to know that there's a way to do inference even in tough cases. 



## Problem 2 - Homicides in US Cities

### *Data Import and Cleanning*

In the following code chunk, I did:

* Creating a `city_state` variable with combine the information form `city` and `state`, eg. Baltimore, MD
* Creating a binary variable that indicates whether the homicide is solved (0 = unsolved; 1 = solved)
* Omitting cities "Dallas, TX," "Phoenix, AZ," "Kansas City, MO," and "Tulsa, AL." Their data is not complete or correct.
* Omitting people who is not White or Black
* Changing `victim_age` into numeric variable. "NA" is assignied automatically for unknown ages

```{r}
homicide_df = 
  read_csv("data/homicide-data.csv") %>%
  janitor::clean_names() %>% 
  mutate(city_state = str_c(city, state, sep = ","),
         solve_status = ifelse(disposition == "Closed by arrest", 1, 0)) %>% 
  filter(city_state != "Dallas,TX" 
         & city_state != "Phoenix,AZ" 
         & city_state != "Kansas City,MO"
         & city_state != "Tulsa,AL") %>% 
  filter(victim_race == "White" | victim_race == "Black") %>% 
  mutate(victim_age = as.numeric(victim_age)) %>%
  select(city_state, solve_status, victim_race, victim_age, victim_sex)
```


### *Fit Logistic Regression via `glm` in Baltimore,MD*

Using the `glm` function with the `family=binomial()` specified to account for the non-Gaussian outcome distribution.

In the follwing code, I fitted a logistic regression and obtained the estimate and confidence interval of the adjusted odds ratio for solving homicides comparing male victims to female victims keeping all other variables fixed.

```{r}
baltimore_df = 
  homicide_df %>% 
  filter(city_state == "Baltimore,MD")

fit_logistic = 
  baltimore_df %>% 
  glm(solve_status ~ victim_age + victim_sex + victim_race, data = ., family = binomial()) %>% 
  broom::tidy() %>% 
  mutate(OR = exp(estimate),
         CI_95_low = exp(estimate - (1.96 * std.error)),
         CI_95_up = exp(estimate + (1.96 * std.error))) %>%
  select(term, OR, CI_95_low, CI_95_up) %>% 
  filter(term == "victim_sexMale") %>% 
  knitr::kable(digits = 3)
```


### *Fit Logistic Regression via `glm` in all cities*

Now run `glm` for each of the cities in your dataset, and extract the adjusted odds ratio (and CI) for solving homicides comparing male victims to female victims. Do this within a “tidy” pipeline, making use of purrr::map, list columns, and unnest as necessary to create a dataframe with estimated ORs and CIs for each city.

```{r}
homicide_nest = nest(homicide_df, data = solve_status:victim_sex) %>% 
  mutate(
    models = map(data, ~ glm(solve_status ~ victim_age + victim_sex + victim_race, data = ., family = binomial())),
    results = map(models, broom::tidy)) %>% 
  unnest(results) %>% 
  mutate(OR = exp(estimate),
         CI_95_low = exp(estimate - (1.96 * std.error)),
         CI_95_up = exp(estimate + (1.96 * std.error))) %>%
  select(city_state, term, OR, CI_95_low, CI_95_up)

adj_OR_sex =
  homicide_nest %>% 
  filter(term == "victim_sexMale")

adj_OR_sex %>% 
  knitr::kable(digits = 3)
```


### *Plot of Estimated Adjusted ORs and CIs among Cities*

```{r}
plot_1 = 
  adj_OR_sex %>% 
  mutate(city_state = fct_reorder(city_state, OR)) %>% 
  ggplot(aes(x = city_state, y = OR)) + 
  geom_point() + 
  geom_point(data = adj_OR_sex, mapping = (aes(x = city_state, y = CI_95_low, color = "blue"))) +
  geom_point(data = adj_OR_sex, mapping = (aes(x = city_state, y = CI_95_up, color = "red"))) +
  theme(axis.text.x = element_text(angle = 80, hjust = 1)) +
  scale_color_manual(values = c("blue", "red"),
                     labels = c("95% CI lower bound",
                                "95% CI upper bound"))

plot_1
```



## Problem 3 - Child Birthweight

